---
title: "Data Science for Economists"
# subtitle: "<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
subtitle: "A Quick Primer on Auctions"
author: "Alex Marsh"
date: "University of North Carolina | [ECON 370](https://github.com/alexiom/ECON370)" #"`r format(Sys.time(), '%d %B %Y')`"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
name: toc

```{css, echo=FALSE}
@media print {
  .has-continuation {
    display: block !important;
  }
}
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center", 
  fig.height=4, #fig.width=6, 
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=F#, echo=F, warning=F, message=F
  )
library(fontawesome)
library(microbenchmark)
library(data.table)
library(ggplot2)
set.seed(123)
```

# Table of contents

1. [Auctions](#auctions)

---
class: inverse, center, middle
name: auctions

# Auctions

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

---
# Motivation

Auction theory has been one of the successes of economics in the past 30-40 years. 

Since auctions are very common in the real world, economists who study empirical auctions have also cropped up to estimate the models the theorists have developed.

While we will not delve too deeply into auctions, I do need to give you just enough information for the problem set. 

---
# Auction Models

The way we model auctions is that bidders have some *valuation* $v_i$ that comes from some distribution $F_v$. The valuation is how much one "values" the object.

The valuations could be iid<sup>1</sup> or iid conidtional on observables or have correlated valuations. 

.footnote[
<sup>1</sup> Independent and identically distributed. This means each valuation is independent of all others and all valuations come from the same distribution.
]

Bidders then know the rules of the auction and submit bids. They bid in such a way that maximizes their expected utility: $$\max_{b_i} v_i \text{Pr}(\text{win with } b_i) - E[p(b_i) | b_i]$$

I am keeping this vague because there are different auction formats and not all require paying the exact bid you submitted e.g. second price auctions.

---
# Auction Models (Cont.)

Once a bidder has her valuation, we think that she forms some optimal bidding strategy $b(v_i)$ that maps her valuation to a bid. We denote the observed bid $b_i$ and we say that $b_i = b(v_i)$ with equilibrium bidding.

In empirical auctions, the goal is to be able to recover $v_i$ or $F_v$ from a bunch of bidding data $(b_i)_{i=1}^{N}$, information about how the auction was ran, and assumptions about bidding behavior. 

As long as $b(v_i)$ is monotonic (higher the valuation, higher the bid), then recovering $v_i$ comes down to inverting $b(v_i)$, or since $b_i = b(v_i)$, then $v_i = b^{-1}(b_i)$.

Some auction formats: 
1. First-Price Sealed Bid
2. Second-Price Sealed Bid
2. English (Ascending) 
3. Dutch
5. All-Pay

---
# Auction Formats

- First-price Sealed Bid: All bidders submit a secret bid, then highest bidder wins and pays the bid.
- Second-Price Sealed Bid: All bidders submit a secret bid, then highest bidder wins but pays the second highest bid.
- English: Bidding starts at some initial bid and bidders can incrementally increase the bid until only one bidder remains. Sometimes there is a minimum bid increment. 
- Dutch: The price starts at a very large value and incrementally decreases until someone declares they are willing to pay that price for the object.
 - "Strategically equivalent" to first-price sealed bid auction.
- All-Pay: Similar to other sealed bid auctions, but everyone pays their bid regardless of winning or loosing.
 - Sometimes used to model R&D races/competitions.


```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
infile = list.files(pattern = '.html')
pagedown::chrome_print(input = infile, timeout = 100)
```